Determinants of Soccer Players' Market Values in Major European League(s)

### 2024-11-29 paper/paper.qmd
Added some photos for the paper and updated introduction and data part

---
title: "Determinants of Soccer Players' Market Values in Major European Leagues"
subtitle: "A Comparative Linear Regression Analysis Across Five Leagues"
author: 
  - John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Player_Market_Value_Analysis"
date: today
date-format: long
abstract: ""
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(arrow)
library(ggplot2)
library(dplyr)
library(modelsummary)
library(tidyverse)
library(knitr)
library(validate)
library(testthat)
```

# Introduction{#sec-intro}
The global soccer transfer market is a dynamic and high-stakes environment where players’ market values are central to team-building strategies, investments, and negotiations. Understanding what drives these values is a critical question for analysts, clubs, and stakeholders. While common factors such as goals scored or club reputation are widely acknowledged to affect valuations, their relative importance likely varies across leagues with differing levels of competition, financial resources, and playing styles.

This paper focuses on five major European soccer leagues—Premier League(England), La Liga (Spain), Serie A (Italy), Bundesliga (Germany), and Ligue 1 (France). These leagues are not only among the most competitive but also vary in terms of audience size, financial backing, and player scouting pipelines, making them ideal for a comparative study of market value determinants.

Using player data from the 2023/24 season, this study develops league-specific linear regression models with market value as the dependent variable. Key predictors include age, goals, assists, club ranking, national team ranking, minutes played, and position. The analysis reveals both shared and league-specific patterns in the factors influencing market value. For instance, goals and assists consistently emerge as significant predictors, but their impact differs by league. Similarly, younger players tend to command higher valuations universally, with age effects most pronounced in Serie A.

These findings contribute to our understanding of how soccer markets function across different leagues and provide actionable insights for clubs and agents. The structure of this paper is as follows: Section 2 describes the dataset and variables, Section 3 outlines the modeling approach, Section 4 presents the results, and Section 5 discusses the implications and limitations of the findings.

@@

## Dataset Overview

The data includes information on players from the top five European leagues during the 2023/24 season. Variables of interest are:

- **Market Value**: Player’s estimated value in Euros.
- **Age**: Player age at the start of the season.
- **Goals and Assists**: Indicators of offensive contributions.
- **Club Ranking**: Rank of the player's club based on league and international performance.
- **National Team Ranking**: Ranking of the player's national team (1 to 210).
- **Minutes Played**: Total minutes played during the season.
- **Position**: Categorized as Defender (DF), Midfielder (MF), Forward (FW), and hybrids like FWDF.

## England

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-emp
#| fig-cap: Distribution of Market Value (in Million Euros)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-efp
#| fig-cap: Distribution of other Variables

# Reshape data for faceting (excluding market_value)
data_long <- england_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
england_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#D62728", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
england_facet_plot
```

## France

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-fmp
#| fig-cap: Distribution of Market Value (in Million Euros)

@@

# Individual plot for market value
france_market_value_plot <- ggplot(france_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-ffp
#| fig-cap: Distribution of other Variables

# Reshape data for faceting (excluding market_value)
data_long <- france_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
france_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
france_facet_plot
```

## Germany

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-gmp
#| fig-cap: Distribution of Market Value (in Million Euros)

@@

# Individual plot for market value
germany_market_value_plot <- ggplot(germany_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-gfp
#| fig-cap: Distribution of other Variables

# Reshape data for faceting (excluding market_value)
data_long <- germany_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
germany_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
germany_facet_plot
```

## Italy

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-imp
#| fig-cap: Distribution of Market Value (in Million Euros)

@@

# Individual plot for market value
italy_market_value_plot <- ggplot(italy_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#2CA02C", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-ifp
#| fig-cap: Distribution of other Variables

# Reshape data for faceting (excluding market_value)
data_long <- italy_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

>>>>

---
title: "Determinants of Soccer Players' Market Values in Major European League(s)"
subtitle: "A Comparative Linear Regression Analysis Across Five Leagues"
author: 
  - John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Player_Market_Value_Analysis"
date: today
date-format: long
abstract: ""
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(arrow)
library(broom)
library(ggplot2)
library(dplyr)
library(modelsummary)
library(tidyverse)
library(knitr)
library(validate)
library(testthat)
library(styler)
library(caret)
library(kableExtra)
```

# Introduction{#sec-intro}
Certainly! Here's an updated introduction that explicitly mentions all the packages from the references you provided and integrates their relevance to the study:

The global soccer transfer market is a high-stakes, multi-billion-Euro industry where decisions on player valuations can significantly influence team-building strategies, investment decisions, and negotiations. Player market values are determined by a mix of tangible performance metrics, such as goals scored or assists provided, and intangible factors like potential, reputation, and team context. While certain predictors are universally recognized, such as offensive contributions and age, the relative importance of these factors varies widely across leagues with differing financial capabilities, tactical preferences, and competitive levels. Understanding the dynamics behind these variations is crucial for clubs, agents, and analysts aiming to optimize their strategies in this complex market.

This study examines the determinants of soccer players’ market values across five major European leagues: the English Premier League, La Liga (Spain), Serie A (Italy), Bundesliga (Germany), and Ligue 1 (France). These leagues not only dominate the global football landscape but also present unique economic and competitive ecosystems, making them ideal for a comparative analysis. Using player data from the 2023/24 season, sourced from platforms such as **Transfermarkt**, **Stathead**, **FIFA.com**, and **Footballdatabase.com**, this research addresses a critical gap: how do the drivers of player market value differ across these leagues?

To conduct this analysis, the study employs league-specific linear regression models with player market value as the dependent variable. Key predictors include performance metrics (e.g., goals, assists, and minutes played), contextual factors (e.g., club and national team rankings), and demographic variables (e.g., age and position). The study relies heavily on the statistical computing environment **R** and the programming language **Python**, employing a suite of packages to ensure methodological rigor and transparency:

- **Data Retrieval and Cleaning**: The Python library `BeautifulSoup` was used to scrape market value and player performance data from Transfermarkt and other sources. In R, the `arrow` package facilitated seamless data reading and processing, while `validate` and `testthat` ensured the integrity and accuracy of the dataset.
- **Data Manipulation and Summarization**: The `tidyverse` collection, including `dplyr` for data manipulation and `ggplot2` for data visualization, formed the backbone of the analysis. Additional tools like `styler` enhanced code readability and reproducibility.
- **Modeling and Summaries**: Regression models were built using the `caret` package for efficient training and testing, while `modelsummary` provided clear, publication-ready tables summarizing model outputs.
- **Dynamic Report Generation**: The `knitr` and `kableExtra` packages were integral to generating dynamic, visually appealing reports that integrated results, tables, and figures.
- **Visualization**: Advanced graphics, including histograms and bar charts, were created with `ggplot2`, showcasing the distribution of market values and key predictors.
- **Statistical Validation**: The robustness of the analysis was supported by packages like `arrow` for handling large datasets and `validate` for data validation.

The results reveal both universal and league-specific patterns in player valuation. Offensive contributions, such as goals and assists, consistently emerged as significant predictors, with goals having the strongest influence in the Premier League and assists being most impactful in the Bundesliga. Age also played a critical role, with younger players commanding higher valuations, particularly in La Liga. Contextual variables like club and national team rankings demonstrated notable variation, reflecting the differing emphasis on team prestige across leagues.

This study contributes to the literature on soccer economics by uncovering how market dynamics vary across Europe’s top leagues. For clubs, the findings provide actionable insights for tailoring recruitment and investment strategies. For agents, the results highlight the most marketable player attributes for negotiations. Methodologically, the integration of advanced statistical and programming tools offers a replicable framework for analyzing market dynamics in soccer and other sports.

The paper is structured as follows: Section 2 describes the dataset and variable construction processes, emphasizing the integration of data from multiple sources. Section 3 details the modeling approach and validation techniques. Section 4 presents the empirical results, identifying key differences across leagues. Section 5 discusses the implications of these findings, highlights the limitations of the study, and outlines future research directions. By bridging methodological rigor with actionable insights, this study advances our understanding of the determinants of soccer players’ market values.

@@

The data used for this study is gathered four primary sources: **Transfermarkt.com**, **Stathead.com**, and **FIFA.com** and **Footballdatabase.com**, which captures player performance, demographic, market valuation and club ranking/national team ranking data. This section details the dataset, the process of constructing the final variables, and the broader context of the data used in the analysis.

## Data Sources and Construction

The dataset was created by merging multiple data sources using player names as unique identifiers. **Transfermarkt.com** provided player market values in Euros, which serve as the dependent variable in this study. Performance metrics such as goals and assists were sourced from **Stathead.com**, a platform that aggregates comprehensive statistics on football players. To ensure consistency and comparability, rankings were applied to national teams and clubs. National team rankings (1 to 210) were extracted from **FIFA.com**, while club rankings were sourced from **Footballdatabase.com**, reflecting both domestic and international performances.

After merging these sources, high-level cleaning was conducted to address missing data and standardize variable formats, and the players who are classified as goalkeepers are dropped. The final cleaned datasets for each league include seven variables, as outlined below:

- **Market Value**: The player’s estimated market value, measured in Euros, as provided by Transfermarkt. This serves as the target variable for the analysis.
- **Age**: The player’s age at the start of the season, reflecting their stage in the career lifecycle.
- **Goals and Assists**: Performance metrics capturing offensive contributions, key determinants of player valuation.
- **Club Ranking**: The global ranking of the player’s club, reflecting both domestic and international performance.
- **National Team Ranking**: The ranking of the player’s national team, highlighting their exposure and representation at the international level.
- **Minutes Played**: The total minutes played by the player during the season, used as a proxy for fitness and importance to the team.
- **Position**: Players are categorized into one of six roles: Defender (DF), Forward (FW), Midfielder (MF), and hybrids such as FWDF, FWMF, and MFDF.

## Data Summary

### England Dataset

The England dataset provides a comprehensive view of player characteristics in the English Premier League. As illustrated in @fig-emp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower market value brackets (below €50 million). This skewness reflects a superstar effect, where a small subset of players, often international stars or highly sought-after talent, command disproportionately higher valuations.

@fig-efp explores the distributions of key predictor variables, including age, goals, assists, national team ranking, club ranking, and minutes played. These distributions reveal important patterns in the dataset:

- **Age**: The age distribution peaks around 24-25 years, indicating that the league predominantly features players in their athletic prime. The range extends from 16 years (likely youth players) to 38 years for veteran professionals.
- **Goals and Assists**: Both offensive contributions exhibit a heavily right-skewed pattern, with most players scoring fewer than 5 goals or assists in a season. This suggests that only a minority of players serve as primary offensive contributors.
- **Club Ranking**: Club ranking is inversely related to club performance, with lower values indicating higher-performing teams. The wide range (from 1 to 838) reflects the diversity of clubs in the dataset, from global giants to smaller teams.
- **National Team Ranking**: Similarly, the national team ranking varies significantly, from top-tier nations to lower-ranked teams. Players from lower-ranked national teams may reflect either emerging talent or under-scouted regions.
- **Minutes Played**: Minutes played are distributed across a broad spectrum, with many players logging significant game time (over 1,000 minutes), indicative of their importance to team dynamics.

@fig-epp focuses on the percentage distribution of player positions. Defenders (DF) account for the largest share (32.6%), followed by midfielders (MF) and midfield-forward hybrids (FWMF). Pure forwards (FW) make up 16.4% of the players, with hybrid positions like FWDF forming a smaller subset. This distribution aligns with the league's tactical emphasis on strong defensive and midfield structures, with forwards occupying more specialized roles.

@tbl-se provides a detailed statistical summary of the numeric variables in the dataset. The mean market value of players is €20.99 million, but the median is much lower at €14.00 million, reaffirming the skewness of the market value distribution. Other variables, such as assists and goals, also display low averages relative to their maximum values, highlighting the rarity of elite-level contributions.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-emp
#| fig-cap: Distribution of Market Value of the England Dataset (in Million Euros)
#| fig-pos: "H"

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-efp
#| fig-cap: Distribution of Predictor Variables of the England Dataset
#| fig-pos: "H"

# Reshape data for faceting (excluding market_value)
data_long <- england_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(
    Variable = factor(Variable, levels = c(
      "age",
      "goals",
      "assists",
      "national_team_ranking",
      "club_ranking",
      "minutes_played"
    ))
  )

# Facet plot for the other variables
england_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#D62728", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

england_facet_plot

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-epp
#| fig-cap: Percentage Distribution of Player Positions of the England Dataset
#| fig-pos: "H"

# Calculate percentages for positions
england_position_percentage <- england_data %>%
  count(position) %>%
  mutate(percentage = n / sum(n) * 100)

# Define a custom color palette for positions
position_colors <- c(
  "DF" = "#1E90FF",   # Dark Blue
  "FW" = "#FF4500",   # Bright Red
  "FWDF" = "#FFD700", # Gold
  "FWMF" = "#20B2AA", # Teal
  "MF" = "#32CD32",   # Bright Green
  "MFDF" = "#9370DB"  # Purple
)

# Plot percentage bar chart
england_position_plot <- ggplot(england_position_percentage, aes(x = position, y = percentage, fill = position)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = position_colors) +
  theme_minimal() +
  labs(
    x = "Position",
    y = "Percentage (%)"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5, size = 3) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

england_position_plot

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-se
#| tbl-cap: Summary Statistics of the England Dataset
#| tbl-pos: "H"

# Function to summarize numeric variables
summarize_numeric <- function(data) {
  data %>%
    summarise(
      Mean = sprintf("%.2f", mean(Value, na.rm = TRUE)),
      Median = sprintf("%.2f", median(Value, na.rm = TRUE)),
      SD = sprintf("%.2f", sd(Value, na.rm = TRUE)),
      Min = sprintf("%.2f", min(Value, na.rm = TRUE)),
      Max = sprintf("%.2f", max(Value, na.rm = TRUE))
    )
}

# Summarize numeric variables
numeric_summary <- england_data %>%
  select(
    market_value_million, 
    age, 
    goals, 
    assists, 
    club_ranking, 
    national_team_ranking, 
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

# Print numeric summary
kable(numeric_summary, align = "c") %>%
  kable_styling(full_width = FALSE)

### France Dataset

The France dataset offers insights into player profiles from Ligue 1, one of Europe’s top football leagues. Similar to other leagues, the distribution of market values in the French dataset, as depicted in @fig-fmp, is heavily right-skewed. Most players have market values below €50 million, with only a few elite players commanding significantly higher valuations. This reflects the trend where top-tier talents, often representing national teams or excelling in international competitions, are valued much higher than their peers.

@fig-ffp illustrates the distributions of predictor variables in the dataset:

- **Age**: The age distribution in Ligue 1 peaks around 24 years, with a majority of players between 20 and 30 years old. This highlights the league's balance between developing young talent and retaining experienced professionals. The maximum age in the dataset is 39 years, underscoring the presence of veteran players in the league.
- **Goals and Assists**: The distributions for goals and assists are highly right-skewed, with most players contributing fewer than 5 goals or assists in a season. This indicates that offensive production is concentrated among a small group of players, typically forwards or attacking midfielders.
- **Club Ranking**: Club rankings span a wide range, from 11 to 568, reflecting the diversity in club performance levels within the league.
- **National Team Ranking**: National team rankings also show significant variation, with players representing countries ranked in the top 10 as well as those from lower-ranked national teams. This diversity emphasizes Ligue 1's role as a platform for both elite and emerging talent.
- **Minutes Played**: The distribution of minutes played indicates that a significant proportion of players log substantial game time (over 1,000 minutes), with some exceeding 3,000 minutes, indicating their critical role in team strategies.

@fig-fpp examines the percentage distribution of player positions in Ligue 1. The largest proportion of players are defenders (DF) at 29.1%, followed by midfielders (MF) at 22.3% and midfield-forward hybrids (FWMF) at 24.3%. Forwards (FW) constitute a smaller proportion (13.1%), while hybrid positions such as FWDF and MFDF account for the remaining distribution. This positional breakdown reflects a balanced tactical structure, with a strong emphasis on defensive and midfield roles.

@tbl-sf provides a comprehensive statistical summary of the numeric variables in the dataset. The mean market value stands at €7.98 million, significantly lower than the Premier League, indicating the varying financial dynamics across leagues. The average age is 23.91 years, reinforcing the league's focus on younger talent. Other variables such as goals and assists demonstrate low means relative to their maximum values, reflecting the rarity of exceptional performances.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-fmp
#| fig-cap: Distribution of Market Value of the France Dataset (in Million Euros)
#| fig-pos: "H"

@@

# Individual plot for market value
france_market_value_plot <- ggplot(france_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-ffp
#| fig-cap: Distribution of Predictor Variables of the France Dataset

# Reshape data for faceting and specify the order of variables
data_long <- france_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

# Facet plot for the variables
france_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

france_facet_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-fpp
#| fig-cap: Percentage Distribution of Player Positions of the France Dataset
#| fig-pos: "H"

# Calculate percentages for positions
france_position_percentage <- france_data %>%
  count(position) %>%
  mutate(percentage = n / sum(n) * 100)

# Define a custom color palette for positions
position_colors <- c(
  "DF" = "#1E90FF",   # Dark Blue
  "FW" = "#FF4500",   # Bright Red
  "FWDF" = "#FFD700", # Gold
  "FWMF" = "#20B2AA", # Teal
  "MF" = "#32CD32",   # Bright Green
  "MFDF" = "#9370DB"  # Purple
)

# Plot percentage bar chart
france_position_plot <- ggplot(france_position_percentage, aes(x = position, y = percentage, fill = position)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = position_colors) +
  theme_minimal() +
  labs(
    x = "Position",
    y = "Percentage (%)"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5, size = 3) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
  )

france_position_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-sf
#| tbl-cap: Summary Statistics of the France Dataset
#| fig-pos: "H"

# Summarize numeric variables for France
numeric_summary_france <- france_data %>%
  select(
    market_value_million, 
    age, 
    goals, 
    assists, 
    club_ranking, 
    national_team_ranking, 
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

# Print numeric summary for France
kable(numeric_summary_france, align = "c") %>%
  kable_styling(full_width = FALSE)
```

### Germany Dataset

The Germany dataset captures player characteristics across teams in the Bundesliga. As illustrated in @fig-gmp, the market value distribution is heavily right-skewed, with the majority of players concentrated in the lower market value brackets (below €50 million). This skewness reflects a superstar effect, with a small subset of players driving disproportionately higher market valuations.

@fig-gfp explores the distributions of predictor variables, shedding light on key patterns:

- **Age**: The age distribution peaks around 25 years, indicating that the league features a significant concentration of players in their athletic prime. The range extends from 17 years (likely youth players) to 39 years for veteran professionals.
- **Goals and Assists**: Both offensive contributions display a heavily right-skewed pattern, with most players contributing fewer than 5 goals or assists. This highlights the rarity of elite offensive contributors in the league.
- **Club Ranking**: Club ranking values range from 5 to 750, reflecting the diverse composition of the league, from globally dominant clubs to lower-performing teams.
- **National Team Ranking**: National team rankings show significant variation, with players representing both top-ranked and lower-ranked nations, capturing the Bundesliga's international player diversity.
- **Minutes Played**: The distribution of minutes played indicates substantial variability, with many players exceeding 1,000 minutes. This suggests that a majority of players are regular starters or receive substantial playing time.

The position distribution, depicted in @fig-gpp, reveals that defenders (DF) comprise the largest share of players (31.7%), followed by midfielders (MF) and forward-midfield hybrids (FWMF). Pure forwards (FW) form 9.6% of the players, while hybrid positions like FWDF make smaller contributions. This composition underscores the league's focus on tactical balance, with strong emphasis on midfield and defensive strategies.

@tbl-sg provides a detailed statistical summary of the numeric variables in the dataset. The mean market value is €10.20 million, while the median is much lower at €4.00 million, highlighting the skewed distribution. Similar trends are observed for assists and goals, where the mean is low compared to the maximum values, further emphasizing the role of standout performers in shaping league outcomes.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-gmp
#| fig-cap: Distribution of Market Value of the Germany Dataset (in Million Euros)
#| fig-pos: "H"

@@

# Individual plot for market value
germany_market_value_plot <- ggplot(germany_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  theme_minimal() +
  labs(
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-gfp
#| fig-cap: Distribution of Predictor Variables of the Germany Dataset
#| fig-pos: "H"

# Reshape data for faceting (excluding market_value)
data_long <- germany_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

# Facet plot for the other variables
germany_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
germany_facet_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-gpp
#| fig-cap: Percentage Distribution of Player Positions of the Germany Dataset
#| fig-pos: "H"

# Calculate percentages for positions
germany_position_percentage <- germany_data %>%
  count(position) %>%
  mutate(percentage = n / sum(n) * 100)

# Define a custom color palette for positions
position_colors <- c(
  "DF" = "#1E90FF",   # Dark Blue
  "FW" = "#FF4500",   # Bright Red
  "FWDF" = "#FFD700", # Gold
  "FWMF" = "#20B2AA", # Teal
  "MF" = "#32CD32",   # Bright Green
  "MFDF" = "#9370DB"  # Purple
)

# Plot percentage bar chart
germany_position_plot <- ggplot(germany_position_percentage, aes(x = position, y = percentage, fill = position)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = position_colors) +
  theme_minimal() +
  labs(
    x = "Position",
    y = "Percentage (%)"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5, size = 3) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
  )

germany_position_plot
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-sg
#| tbl-cap: Summary Statistics of the Germany Dataset
#| fig-pos: "H"

# Summarize numeric variables for Germany
numeric_summary_germany <- germany_data %>%
  select(
    market_value_million, 
    age, 
    goals, 
    assists, 
    club_ranking, 
    national_team_ranking, 
    minutes_played
  ) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize_numeric()

# Print numeric summary for Germany
kable(numeric_summary_germany, align = "c") %>%
  kable_styling(full_width = FALSE)
```

### Italy Dataset

The Italy dataset provides an insightful analysis of player characteristics within the Serie A league. As shown in Figure @fig-imp, the market value distribution is right-skewed, with a majority of players concentrated in lower valuation brackets (below €30 million). This distribution is characteristic of the league, where the presence of a few superstar players significantly elevates the upper tail.

Figure @fig-ifp highlights the distributions of key predictor variables:

- **Age**: The age distribution centers around 25 years, reflecting a mix of players in their early careers and those in their prime. The range extends from 15 to 36 years.
- **Goals and Assists**: These variables exhibit heavy skewness, with most players scoring fewer than 5 goals or assists in a season, underscoring the prominence of specialized attacking roles.
- **Club Ranking**: With values ranging from 3 to 681, the club ranking illustrates the diverse spectrum of team performance within Serie A.
- **National Team Ranking**: The national team ranking shows a range from 1 to 141, capturing the representation of players from both elite and emerging national teams.
- **Minutes Played**: The distribution of minutes played indicates significant variability, with several players logging substantial game time (over 1,000 minutes) and a few recording minimal participation.

Figure @fig-ipp depicts the percentage distribution of player positions. Defenders (DF) form the largest proportion (35.4%), followed by midfielders (MF) at 24%. Forward-midfield hybrids (FWMF) and forwards (FW) are well-represented, emphasizing the league's balanced approach to attacking and defensive strategies. Hybrid positions such as FWDF and MFDF form smaller shares of the dataset.

Table @tbl-si presents summary statistics for the numeric variables in the Italy dataset. The mean market value is €9.37 million, with a median of €4.00 million, demonstrating a significant skew. Goals and assists have low averages relative to their maximum values, aligning with the skewed distributions observed in their histograms.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-imp
#| fig-cap: Distribution of Market Value of the Italy Dataset (in Million Euros)
#| fig-pos: "H"

@@

# Individual plot for market value
italy_market_value_plot <- ggplot(italy_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#2CA02C", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

@@

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-ifp
#| fig-cap: Distribution of Predictor Variables of the Italy Dataset
#| fig-pos: "H"

# Reshape data for faceting and specify the order of variables
data_long <- italy_data %>%
  select(
    age,
    goals,
    assists,
    national_team_ranking,
    club_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  mutate(Variable = factor(Variable, levels = c(
    "age", "goals", "assists", "national_team_ranking", "club_ranking", "minutes_played"
  )))

# Facet plot for the variables
italy_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#2CA02C", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

italy_facet_plot

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-ipp
#| fig-cap: Percentage Distribution of Player Positions of the Italy Dataset
#| fig-pos: "H"

# Calculate percentages for positions
italy_position_percentage <- italy_data %>%
  count(position) %>%
  mutate(percentage = n / sum(n) * 100)
